---
title: "Richness of 18S, COI, and 12S"
author: "Anni Djurhuus"
date: "15 August 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r read biom file 18S and 12S, include=FALSE}
require(RColorBrewer)
brewer.pal(9,"Set1")
display.brewer.all()
mypalette <- brewer.pal(9,"Set1")
cbbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
#biocLite("phyloseq")
library(phyloseq)
#16S
#m2w_16S <- import_biom("/Users/anni_djurhuus/Documents/Projects/MBON/m2w/16S/")
#m2w_16S <- subset_samples(m2w_16S, sample_type == "environmental")
#metadata_16S <- m2w_16S@sam_data
#otu_16S <- otu_table(m2w_16S, taxa_are_rows=TRUE)

#18S
m2w_18S <- import_biom("/Users/anni_djurhuus/Documents/Projects/MBON/m2w/18S/MB/M2W_18S_80per_NT_json_obs_md.biom")
m2w_18S <- subset_samples(m2w_18S, sample_type == "environmental")
m2w_18S <- subset_samples(m2w_18S, sample_names(m2w_18S) !="10706C1_0m_1_S51" & sample_names(m2w_18S) !="10706C1_0m_2_S52" & sample_names(m2w_18S) !="10706C1_0m_3_S53" & sample_names(m2w_18S) !="15806C1_0m_1_S54" & sample_names(m2w_18S) !="15806C1_0m_2_S55" & sample_names(m2w_18S) !="15806C1_0m_3_S56" & sample_names(m2w_18S) !="27006C1_0m_1_S57" & sample_names(m2w_18S) !="27006C1_0m_2_S58" & sample_names(m2w_18S) !="27006C1_0m_3_S59" & sample_names(m2w_18S) !="33306C1_0m_1_S60" & sample_names(m2w_18S) !="33306C1_0m_2_S61" & sample_names(m2w_18S) !="33306C1_0m_3_S62")
metadata_18S <- m2w_18S@sam_data
library(stringr)
metadata_18S2 <- str_split_fixed(as.character(metadata_18S$DATE_TIME_GMT),' ', 2)
colnames(metadata_18S2) <- c("YEAR_DATE","TIME")
metadata_18S2b <- str_split_fixed(as.character(metadata_18S2[,1]),'-', 3)
colnames(metadata_18S2b) <- c("year","month","day")
season <- c("Spring","Spring","Spring","Summer","Summer","Summer","Autumn","Autumn","Autumn","Winter","Winter","Winter","Spring","Spring","Spring","Summer","Summer","Summer","Autumn","Autumn","Autumn","Autumn","Autumn","Autumn","Spring","Spring","Spring","Summer","Summer","Summer","Autumn","Autumn","Autumn","Winter","Winter","Winter","Spring","Spring","Spring","Summer","Summer","Summer","Autumn","Autumn","Autumn","Winter","Winter","Winter")
metadata_18S <- sample_data(cbind(season,metadata_18S,metadata_18S2, metadata_18S2b))

otu_18S <- otu_table(m2w_18S, taxa_are_rows=TRUE)
b <- as.data.frame(otu_18S)
#tax_18S <- as.data.frame(tax_table(m2w_18S))
m2w_18S <- merge_phyloseq(metadata_18S,otu_18S)

#12S
m2w_COI <- import_biom("/Users/anni_djurhuus/Documents/Projects/MBON/m2w/COI/COI_m2w_twostep_json_obs_md.biom")
m2w_COI <- subset_samples(m2w_COI, sample_type == "environmental")
m2w_COI <- subset_samples(m2w_COI, sample_names(m2w_COI) !="JD10706c1_0m_1" & sample_names(m2w_COI) !="JD10706c1_0m_2" & sample_names(m2w_COI) !="JD10706c1_0m_3" & sample_names(m2w_COI) !="JD15806c1_0m_1" & sample_names(m2w_COI) !="JD15806c1_0m_2" & sample_names(m2w_COI) !="JD15806c1_0m_3" & sample_names(m2w_COI) !="JD27006c1_0m_1" & sample_names(m2w_COI) !="JD27006c1_0m_2" & sample_names(m2w_COI) !="JD27006c1_0m_3" & sample_names(m2w_COI) !="JD33306c1_0m_1" & sample_names(m2w_COI) !="JD33306c1_0m_2" & sample_names(m2w_COI) !="JD33306c1_0m_3")
m2w_COI <- subset_samples(m2w_COI, DEPTH_M == 0)
metadata_COI <- m2w_COI@sam_data
otu_COI <- otu_table(m2w_COI, taxa_are_rows=TRUE)
tax_COI <- as.data.frame(tax_table(m2w_COI))
#12S
m2w_12S <- import_biom("/Users/anni_djurhuus/Documents/Projects/MBON/m2w/12S/MB_12S_NT_json_obs_md.biom")
m2w_12S <- subset_samples(m2w_12S, sample_names(m2w_12S) != "ArtComm_HS_1" & sample_names(m2w_12S) != "ArtComm_HS_2" & sample_names(m2w_12S) !="ArtComm_HS_3" & sample_names(m2w_12S) !="CB_CANON160925_1" & sample_names(m2w_12S) !="CB_CANON160925_2" & sample_names(m2w_12S) !="CB_CANON160925_3" & sample_names(m2w_12S) !="EB_11162016" & sample_names(m2w_12S) !="EB_11212016" & sample_names(m2w_12S) !="EB_12282016" & sample_names(m2w_12S) !="NTC_pool_1" & sample_names(m2w_12S) !="NTC_pool_2" & sample_names(m2w_12S) !="NTC_pool_3" & sample_names(m2w_12S) !="JD10706c1_0m_1" & sample_names(m2w_12S) !="JD10706c1_0m_2" & sample_names(m2w_12S) !="JD10706c1_0m_3" & sample_names(m2w_12S) !="JD15806c1_0m_1" & sample_names(m2w_12S) !="JD15806c1_0m_2" & sample_names(m2w_12S) !="JD15806c1_0m_3" & sample_names(m2w_12S) !="JD27006c1_0m_1" & sample_names(m2w_12S) !="JD27006c1_0m_2" & sample_names(m2w_12S) !="JD27006c1_0m_3")
#m2w_12S <- subset_samples(m2w_12S, sample_type == "environmental")
otu_12S <- otu_table(m2w_12S, taxa_are_rows=TRUE)
metadata_12S <- m2w_12S@sam_data
#names(metadata_12S)
```


```{r Plot richness for 12S and 18S,echo=FALSE, include=FALSE}
#plot richness for 12S, COI, and 18S of raw sequences
#install.packages("cowplot")
require(cowplot)
library(ggplot2)
#p <- plot_richness(m2w_16S,x="DATE_TIME_GMT",title="16S",measures = "Observed") + theme(axis.title.x=element_blank())
#p <- p + geom_point(size=5, alpha=0.4,position=position_jitter(width=0.1))
#p <- p + geom_boxplot(outlier.color = NA, fill=NA)
#p + geom_point(size=5, alpha=0.4)

p1 <- plot_richness(m2w_18S,x="YEAR_DATE",title="18S",measures = "Observed") + theme(axis.title.x=element_blank()) 
#p1 <- p1 + geom_point(size=5, alpha=0.4,position=position_jitter(width=0.1))
p1 <- p1 + geom_boxplot(outlier.color = NA, fill=NA)
p1

p2 <- plot_richness(m2w_COI,x="DATE_TIME_GMT",title="COI",measures = "Observed") + theme(axis.title.x=element_blank())
#p2 <- p2 + geom_point(size=5, alpha=0.4,position=position_jitter(width=0.1))
p2 <- p2 + geom_boxplot(outlier.color = NA, fill=NA)
p2 + geom_point(size=5, alpha=0.4)

p3 <- plot_richness(m2w_12S,x="DATE_TIME_GMT",title="12S",measures = "Observed") + theme(axis.title.x=element_blank())
#p3 <- p3 + geom_point(size=5, alpha=0.4,position=position_jitter(width=0.1))
p3 <- p3 + geom_boxplot(outlier.color = NA, fill=NA)
p3 + geom_point(size=5, alpha=0.4)
#plot_grid(p1, p2, p3, labels = c("A", "B", "C"), align="h")
```


```{r rarefy sequences, include=FALSE}
#colSums(otu_16S)
colSums(otu_18S)
colSums(otu_COI)
colSums(otu_12S)

#rare_16S = rarefy_even_depth(m2w_16S)
rare_18S = rarefy_even_depth(m2w_18S)
rare_COI = rarefy_even_depth(m2w_COI)
rare_12S = rarefy_even_depth(m2w_12S)
```

```{r Plot richness for 12S and 18S post rarefication,echo=FALSE, include=FALSE}
#plots of richness post rarefaction
#install.packages("cowplot")
require(cowplot)
#p4 <- plot_richness(rare_16S,x="DATE_TIME_GMT",title="16S",measures = "Observed") + theme(axis.title.x=element_blank())
#p4 <- p4 + geom_point(size=5, alpha=0.4,position=position_jitter(width=0.1))
#p4 <- p4 + geom_boxplot(outlier.color = NA, fill=NA)
#p4 + geom_point(size=5, alpha=0.4)

p5 <- plot_richness(rare_18S,x="YEAR_DATE",title="18S",measures = "Observed") + theme(axis.title.x=element_blank())
#p5 <- p5 + geom_point(size=5, alpha=0.4,position=position_jitter(width=0.1))
p5 <- p5 + geom_boxplot(outlier.color = NA, fill=NA)
p5

p6 <- plot_richness(rare_COI,x="DATE_TIME_GMT",title="COI",measures = "Observed") + theme(axis.title.x=element_blank())
#p6 <- p6 + geom_point(size=5, alpha=0.4,position=position_jitter(width=0.1))
p6 <- p6 + geom_boxplot(outlier.color = NA, fill=NA)
p6 + geom_point(size=5, alpha=0.4)

p7 <- plot_richness(rare_12S,x="DATE_TIME_GMT",title="12S",measures = "Observed") + theme(axis.title.x=element_blank())
#p7 <- p7 + geom_point(size=5, alpha=0.4,position=position_jitter(width=0.1))
p7 <- p7 + geom_boxplot(outlier.color = NA, fill=NA)
p7 + geom_point(size=5, alpha=0.4)
```


```{r DESeq2, eval=FALSE, include=FALSE}
#DESeq2
library(DESeq2)
#16S
#deseq_18s = phyloseq_to_deseq2(m2w_18S, ~ DATE_TIME_GMT)
#deseq_18s = DESeq(deseq_18s, test="Wald", fitType="parametric")
#deseq_18s = estimateSizeFactors(deseq_18s)
#deseq_18s <- counts(deseq_18s, normalized=TRUE)
#m2w_18S2 <- otu_table(deseq_18s, taxa_are_rows = TRUE)

#18S
deseq_18s = phyloseq_to_deseq2(m2w_18S, ~ DATE_TIME_GMT)
deseq_18s = DESeq(deseq_18s, test="Wald", fitType="parametric")

res = results(deseq_18s, cooksCutoff = FALSE)
alpha = 0.01
sigtab = res[which(res$padj < alpha), ]
sigtab = cbind(as(sigtab, "data.frame"), as(tax_table(m2w_18S)[rownames(sigtab), ], "matrix"))
head(sigtab)

library("ggplot2")
theme_set(theme_bw())
scale_fill_discrete <- function(palname = "Set1", ...) {
    scale_fill_brewer(palette = palname, ...)
}
# Phylum order
x = tapply(sigtab$log2FoldChange, sigtab$Rank1, function(x) max(x))
x = sort(x, TRUE)
sigtab$Rank1 = factor(as.character(sigtab$Rank1), levels=names(x))
# Genus order
x = tapply(sigtab$log2FoldChange, sigtab$Rank6, function(x) max(x))
x = sort(x, TRUE)
sigtab$Rank6 = factor(as.character(sigtab$Rank6), levels=names(x))
ggplot(sigtab, aes(x=Rank6, y=log2FoldChange, color=Rank1)) + geom_point(size=6) + 
  theme(axis.text.x = element_text(angle = -90, hjust = 0, vjust=0.5))

deseq_18s = estimateSizeFactors(deseq_18s)
deseq_18s <- counts(deseq_18s, normalized=TRUE)
a <- as.data.frame(deseq_18s)
otu2_18S <- otu_table(deseq_18s, taxa_are_rows = TRUE)
m2w_18S2 <- merge_phyloseq(metadata_18S,otu2_18S,tax_18S)

#COI
#deseq_COI = phyloseq_to_deseq2(m2w_COI, ~ DATE_TIME_GMT)
# deseq_COI = DESeq(deseq_COI, test="Wald", fitType="parametric")
# deseq_COI = estimateSizeFactors(deseq_COI)
# deseq_COI <- counts(deseq_COI, normalized=TRUE)
# m2w_18S2 <- otu_table(deseq_COI, taxa_are_rows = TRUE)
# 
# #12S
# #18S
# deseq_12s = phyloseq_to_deseq2(m2w_12S, ~ DATE_TIME_GMT)
# deseq_12s = DESeq(deseq_12s, test="Wald", fitType="parametric")
# deseq_12s = estimateSizeFactors(deseq_12s)
# deseq_12s <- counts(deseq_12s, normalized=TRUE)
# m2w_12S2 <- otu_table(deseq_12s, taxa_are_rows = TRUE)

```


```{r Richness post rarefaction and occupancy modelling,echo=FALSE, include=FALSE}
#Observed richness post rarefaction and occupancy modelling
#install.packages("lubridate")
library(lubridate)

#18S
occu_18S=read.csv("/Users/anni_djurhuus/Documents/Projects/MBON/m2w/18S/MB/Occ_Model_18s_taxa2.csv",header=TRUE, row.names=1) #12S
occu_18S <- occu_18S[,8:23]
occu_18S <- t(occu_18S)
meta_18S_occup <- m2w_18S@sam_data
b <- metadata_18S[seq(1, length(metadata_18S), 3)]
b <- b[1:16,]
season2 <- c("Spring","Summer","Autumn","Winter","Spring","Summer","Autumn","Autumn","Spring","Summer","Autumn","Winter","Spring","Summer","Autumn","Winter")
binsum=apply(occu_18S, 1, sum)
#Append row sums to binary table
datasheet_binary = cbind(occu_18S, binsum)
datasheet_binary = cbind(b,season2, datasheet_binary)
#datasheet_binary <- datasheet_binary[1:16,]
datasheet_binary$DATE_TIME_GMT <- ymd_hms(datasheet_binary$DATE_TIME_GMT)

#COI
occu_COI=read.csv("/Users/anni_djurhuus/Documents/Projects/MBON/m2w/COI/Occ_Model_COI_taxa2.csv",header=TRUE, row.names=1) #12S
occu_COI <- occu_COI[,8:23]
occu_COI <- t(occu_COI)

metadata_COI <- m2w_COI@sam_data
c <- metadata_COI[seq(1, length(metadata_COI), 3)]
c <- c[1:16,]
season <- c("Spring","Summer","Autumn","Winter","Spring","Summer","Autumn","Autumn","Spring","Summer","Autumn","Winter","Spring","Summer","Autumn","Winter")
binsum_COI=apply(occu_COI, 1, sum)
#Append row sums to binary table
datasheet_binary_COI = cbind(occu_COI, binsum_COI,season)
datasheet_binary_COI = cbind(c, datasheet_binary_COI)
datasheet_binary_COI <- datasheet_binary_COI[1:16,]
datasheet_binary_COI$DATE_TIME_GMT <- ymd_hms(datasheet_binary_COI$DATE_TIME_GMT)

#12S
# occu_12S=t(read.csv("/Users/anni_djurhuus/Documents/Projects/MBON/m2w/12S/Occ_Model_12S.csv",header=TRUE, row.names=1)) #12S
# occu_12S <- occu_12S[c(-9,-13),]
# metadata_12S <- m2w_12S@sam_data
# d <- metadata_12S[seq(1, length(metadata_12S), 3)]
# d <- d[1:14,]
# 
# binsum_12S=apply(occu_12S, 1, sum)
# #Append row sums to binary table
# datasheet_binary_12S = cbind(occu_12S, binsum_12S)
# datasheet_binary_12S = cbind(d, datasheet_binary_12S)
# datasheet_binary_12S <- datasheet_binary_12S[1:16,]
# datasheet_binary_12S$DATE_TIME_GMT <- ymd_hms(datasheet_binary_12S$DATE_TIME_GMT)

#PLOTS 
plot(datasheet_binary$binsum~datasheet_binary$DATE_TIME_GMT, type="o", pch=16, ylim=c(0,1400),col="#E41A1C", xlab="Date", ylab="Observed richness")
par(new=TRUE)
plot(datasheet_binary_COI$binsum~datasheet_binary_COI$DATE_TIME_GMT, ylim=c(0,1400),type="o", pch=16, col="#377EB8", xlab="Date", ylab="Observed richness")
par(new=TRUE)
#plot(datasheet_binary_12S$binsum~datasheet_binary_12S$DATE_TIME_GMT, ylim=c(0,150),type="o", pch=16, col="black", xlab="Date", ylab="Observed richness")
par(new=TRUE)
plot(datasheet_binary_COI$TMP_C~datasheet_binary_COI$DATE_TIME_GMT,type="l",lwd=2,lty=2, pch=16, col="grey", xlab="Date", ylab="Observed richness")
legend("topleft",legend=c("18S","COI", "12S"), pch=c(16,16,16), col=c("#E41A1C","#377EB8","black"))


lm_temp_18S <- lm(TMP_C~binsum, data=datasheet_binary)
newtemp_18s <- seq(min(datasheet_binary$binsum), max(datasheet_binary$binsum), 0.35)
conf_lm_temp_18S <- predict(lm_temp_18S, newdata=data.frame(binsum=newtemp_18s), interval="confidence")
summary(lm_temp_18S)

pdf(file="/Users/anni_djurhuus/Documents/Projects/MBON/m2w/figs/18S_richness_temperature.pdf",width=5.5,height=7)
plot(datasheet_binary$binsum, datasheet_binary$TMP_C,main="18S richness with temperature", col=c(cbbPalette[2], cbbPalette[3], cbbPalette[4], cbbPalette[5])[as.factor(datasheet_binary$season)], pch=c(16,17,15,18)[as.factor(datasheet_binary$year)],cex=1.8,xlab="Richness (OTUs)",ylab="Temperature")
abline(lm_temp_18S, col="grey",lty=1)
lines(newtemp_18s,conf_lm_temp_18S[,2], lty=3,col="grey")
lines(newtemp_18s,conf_lm_temp_18S[,3], lty=3,col="grey")
legend("topleft",ncol=2, bty="n",legend=c("Spring", "Summer","Autumn", "Winter","2013","2014","2015","2016"), col=c(cbbPalette[3], cbbPalette[2], cbbPalette[4], cbbPalette[5],"black","black","black","black"), pch=c(16,16,16,16,16,17,15,18))
dev.off()

plot(datasheet_binary_COI$binsum, datasheet_binary_COI$TMP_C,col=c(cbbPalette[2],cbbPalette[3],cbbPalette[4],cbbPalette[5])[as.factor(datasheet_binary$season)],pch=c(16,17,15,18)[as.factor(datasheet_binary$year)],cex=1.4)
#plot(datasheet_binary_12S$binsum, datasheet_binary_12S$TMP_C)
```

```{r richness by taxa post rarefaction plus occupancy modelling,echo=FALSE, include=FALSE}
#Richness by taxa post rarefaction and occupancy modelling
#18S
taxa_18S=read.csv("/Users/anni_djurhuus/Documents/Projects/MBON/m2w/18S/MB/Occ_Model_18S_taxa2.csv",header=TRUE, row.names=1) 
#18S
taxa_grouped_18S <- tax_table(as.matrix(taxa_18S[1:7]))
otu_18S <- otu_table(taxa_18S[8:23], taxa_are_rows=TRUE)
sample_names(otu_18S)
row.names(b) <- sample_names(otu_18S)

taxa_grouped_18S <- merge_phyloseq(taxa_grouped_18S,otu_18S,b)
genus_glom_18S = tax_glom(taxa_grouped_18S, "Genus")
family_glom_18S = tax_glom(taxa_grouped_18S, "Family")

glom_18S_p1 <- plot_richness(genus_glom_18S,x="YEAR_DATE",title="18S Genus",measures = "Observed") + theme(axis.title.x=element_blank())
#glom_18S_p1 <- glom_18S_p1 + geom_point(size=5, alpha=0.4,position=position_jitter(width=0.1))
#glom_18S_p1 + geom_boxplot(outlier.color = NA, fill=NA)
#glom_18S_p1 + geom_point(size=5, alpha=0.4)

fam_glom_18S_p1 <- plot_richness(family_glom_18S,x="YEAR_DATE",title="18S Family",measures = "Observed") + theme(axis.title.x=element_blank())
#fam_glom_18S_p1 <- fam_glom_18S_p1 + geom_point(size=5, alpha=0.4,position=position_jitter(width=0.1))
#fam_glom_18S_p1 + geom_boxplot(outlier.color = NA, fill=NA)
#fam_glom_18S_p1 + geom_point(size=5, alpha=0.4)



# otu_table_18S <- t(as.data.frame(zoo_18S@otu_table))
# rare_meta_18S <- zoo_18S@sam_data
# otu_meta_18S <- cbind(rare_meta_18S,otu_table_18S)
# 
# dist_18S=vegdist(otu_meta_18S[,c(44:ncol(otu_meta_18S))], method="bray")
# clust.res_18S=hclust(dist_18S, method="average")
# plot(clust.res_18S, labels=otu_meta_18S$de_replicate, hang=-1, cex=0.5, main=NULL, xlab="Sample") #or use labels=n if this doesn't work correctly
# 
# #NMDS 18S
# otu_meta_18S_2 <- droplevels(otu_meta_18S)
# nmds_18S <- metaMDS(otu_meta_18S[,c(44:ncol(otu_meta_18S))], distance="bray", trace = FALSE, trymax=100) 
# 
# xbac_18S <- nmds_18S$points[,1]
# ybac_18S <- nmds_18S$points[,2]
# 
# pdf(file="/Users/anni_djurhuus/Documents/Projects/MBON/eDNA_zoo/fig/18S_NMDS.pdf", height=8, width=6)
# plot(xbac_18S, ybac_18S, xlab="Coordinate 1",ylab="Coordinate 2",main="COI", col=c(col2[otu_meta_18S$de_replicate]), pch=c(15,16,17)[otu_meta_18S$type_sample], cex=1.7)
# #text(nmds_18S$points[,1], nmds_18S$points[,2],labels=otu_meta_18S$Station,pos=3,cex=0.5) 
# legend("bottomright",ncol=1, legend=c("PF-eDNA", "eDNA", "T-DNA"), cex=0.9,col=c("#999999"),pch=c(19,15,17))
# beta_18S = betadiver(otu_meta_18S[,c(44:ncol(otu_meta_18S))], "z")
# adonis(beta_18S ~ type_sample, data= otu_meta_18S, method="bray", perm=500) #make 
# legend("topleft", legend=c("PERMANOVA, p=0.001996"),bty="n", cex=0.8)
# dev.off()
# 

#COI
taxa_COI=read.csv("/Users/anni_djurhuus/Documents/Projects/MBON/m2w/COI/Occ_Model_COI_taxa2.csv",header=TRUE, row.names=1) #COI
taxa_grouped_COI <- tax_table(as.matrix(taxa_COI[1:7]))
otu_COI <- otu_table(taxa_COI[8:23], taxa_are_rows=TRUE)
row.names(c) <- sample_names(otu_COI)
taxa_grouped_COI <- merge_phyloseq(taxa_grouped_COI,otu_COI,c)
genus_glom_COI = tax_glom(taxa_grouped_COI, "Genus")
family_glom_COI = tax_glom(taxa_grouped_COI, "Family")

glom_COI_p1 <- plot_richness(genus_glom_COI,x="DATE_TIME_GMT",title="COI Genus",measures = "Observed") + theme(axis.title.x=element_blank())

fam_glom_COI_p1 <- plot_richness(family_glom_COI,x="DATE_TIME_GMT",title="COI Family",measures = "Observed") + theme(axis.title.x=element_blank())

#12S
#taxa_12S=read.csv("/Users/anni_djurhuus/Documents/Projects/MBON/m2w/12S/Occ_Model_12S_taxa.csv",header=TRUE, row.names=1) #12S

#taxa_grouped_12S <- tax_table(as.matrix(taxa_12S[1:7]))
#otu_12S <- otu_table(taxa_12S[8:24], taxa_are_rows=TRUE)
#row.names(d) <- sample_names(otu_12S)
#taxa_grouped_12S <- merge_phyloseq(taxa_grouped_12S,otu_12S,c)
#genus_glom_12S = tax_glom(taxa_grouped_12S, "Genus")
#family_glom_12S = tax_glom(taxa_grouped_12S, "Family")

#glom_12S_p1 <- plot_richness(genus_glom_12S,x="DATE_TIME_GMT",title="12S Genus",measures = "Observed") + theme(axis.title.x=element_blank())
#glom_12S_p1 <- glom_12S_p1 + geom_point(size=5, alpha=0.4,position=position_jitter(width=0.1))
#glom_12S_p1 <- glom_12S_p1 + geom_boxplot(outlier.color = NA, fill=NA)
#glom_12S_p1 + geom_point(size=5, alpha=0.4)

#fam_glom_12S_p1 <- plot_richness(family_glom_12S,x="DATE_TIME_GMT",title="12S Family",measures = "Observed") + theme(axis.title.x=element_blank())
#fam_glom_12S_p1 <- fam_glom_12S_p1 + geom_point(size=5, alpha=0.4,position=position_jitter(width=0.1))
#fam_glom_12S_p1 <- fam_glom_12S_p1 + geom_boxplot(outlier.color = NA, fill=NA)
#fam_glom_12S_p1 + geom_point(size=5, alpha=0.4)
```

###18S
#####18S OTU richness prior to rarefication and occupancy modeling (left) and post rarefication (right)
```{r,echo=FALSE}
library(cowplot)
plot_grid(p1, p5)
```

#####18S taxa richness post occupancy modeling on genus (left) and family (right)
```{r,echo=FALSE}
library(cowplot)
plot_grid(glom_18S_p1,fam_glom_18S_p1)
```

#####18S NMDS pre rarification and occupancy modelling 
#######Bray-curtis distance matrix was used. The PERMANOVA showed significant differences between seasons, not years (df=3, p=0.001).
```{r run ordination pre rarefy and occupancy, include=FALSE}
ord_18S <- ordinate(m2w_18S, "NMDS", "bray")
library(vegan)
p_18S <-plot_ordination(m2w_18S, ord_18S, "samples",color="season")
orddata_18S <- p_18S
map.df_18S <- data.frame(sample_data(m2w_18S))
dist_18S_pre <- phyloseq::distance(m2w_18S, method="bray")
adonis(dist_18S_pre ~ year, data = map.df_18S)
```

```{r NMDS pre rarefication and occupancy, echo=FALSE, fig.height=5, fig.width=4}
plot(x=orddata_18S$NMDS1, y=orddata_18S$NMDS2, type="n",main="18S pre occupancy and rarefication", xlim=c(-3,3), ylim=c(-2,3), xlab="NMDS1", ylab="NMDS2")
points(ord_18S, disp="si",cex=1.4, col=c(cbbPalette[2], cbbPalette[3], cbbPalette[4], cbbPalette[5])[as.factor(map.df_18S$season)], pch=c(15,16,17,18)[map.df_18S$year])
legend("topleft",ncol=2, bty="n",legend=c("Spring", "Summer","Autumn", "Winter","2013","2014","2015","2016"), col=c(cbbPalette[3], cbbPalette[2], cbbPalette[4], cbbPalette[5],"black","black","black","black"), pch=c(16,16,16,16,15,16,17,18))
legend("bottomright", legend=c("PERMANOVA, p=0.001"),bty="n", cex=0.8)
```

#####18S NMDS post occupancy modelling
#######Bray-curtis distance matrix was used. The PERMANOVA showed significant differences between seasons and years (df=3, p=0.001).
```{r run ordination for 18S post occupancy,include=FALSE}
ord_18S_occu <- ordinate(taxa_grouped_18S, "NMDS", "bray")
library(vegan)
p_18S_occu <-plot_ordination(taxa_grouped_18S, ord_18S_occu, "samples",color="season")
orddata_18S_occu <- p_18S_occu
map.df_18S_occu <- data.frame(sample_data(taxa_grouped_18S))
dist_18S_occu <- phyloseq::distance(taxa_grouped_18S, method="bray")
adonis(dist_18S_occu ~ season, data = map.df_18S_occu)
```

```{r NMDS post occupancy, echo=FALSE, fig.height=5, fig.width=4}
plot(x=orddata_18S_occu$NMDS1, y=orddata_18S_occu$NMDS2,cex=1.4, type="n",main="18S pre occupancy and rarefication", xlim=c(-1.5,1.5), ylim=c(-1.5,1.0), xlab="Axis 1", ylab="Axis 2")
points(ord_18S_occu, cex=1.4,disp="si", col=c(cbbPalette[2], cbbPalette[3], cbbPalette[4], cbbPalette[5])[as.factor(map.df_18S_occu$season)], pch=c(15,16,17,18)[map.df_18S_occu$year])
legend("bottomright",ncol=2, legend=c("Spring", "Summer","Autumn", "Winter","2013","2014","2015","2016"), col=c(cbbPalette[3], cbbPalette[2], cbbPalette[4], cbbPalette[5],"black","black","black","black"), pch=c(16,16,16,16,16,17,15,18))
legend("topleft", legend=c("PERMANOVA, p=0.003"),bty="n", cex=0.8)
```


###COI
#####COI OTU richness prior to rarefication and occupancy modeling (left) and post rarefication (right)
```{r,echo=FALSE}
library(cowplot)
plot_grid(p2, p6)
```

#####COI taxa richness post occupancy modeling on genus (left) and family (right)
```{r,echo=FALSE}
library(cowplot)
plot_grid(glom_COI_p1,fam_glom_COI_p1)
```

###12S
#####12S OTU richness prior to rarefication and occupancy modeling (left) and post rarefication (right)
```{r,echo=FALSE}
#library(cowplot)
#plot_grid(p3, p7)
```

#####12S taxa richness post occupancy modeling on genus (left) and family (right)
```{r,echo=FALSE}
#library(cowplot)
#plot_grid(glom_12S_p1,fam_glom_12S_p1)
```

